---
title: "STAT 545A Assignment 04: Tidy data and joins"
output:
  #pdf_document: default
  html_document: default
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(gapminder)
library(gridExtra)
library(grid)
library(spelling)
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1: Univariate Data Reshaping (30%)

Univariate Option 1:
1. Make a tibble with one row per year, and columns for life expectancy for two or more countries.

 * Consider the Life expectancy data for North America:
 
```{r}
subset <- gapminder %>%
  filter(country %in% c("Canada", "Mexico", "United States")) %>%
  select(country, year, lifeExp) 

subset %>%
   knitr::kable(align = "c", caption = "Life Expectancy Per Year Selected Countries")
```

 * The following table contains the values of Life expectancy in a different column of each country by years in rows:
 
```{r}
(subset_wide <- subset %>% 
  pivot_wider(id_cols = year, 
              names_from = country, 
              values_from = lifeExp))
subset_wide %>%
   knitr::kable(align = "c", caption = "Life Expectancy Per Year Selected Countries")
```




2. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
(subset_wide %>%
  ggplot(aes(Canada, Mexico)) +
  geom_jitter() +
  geom_smooth(method = "loess") +
  geom_point() +
  ggtitle("Canada vs United States life expectancy"))

```

3. Re-lengthen the data.
```{r}
subset_wide %>% 
  pivot_longer(cols = c(-year), 
               names_to  = "country", 
               values_to = "LifeExp") %>%
   knitr::kable(align = "c", caption = "Life Expectancy Per Year Selected Countries")

```



## Exercise 2: Multivariate Data Reshaping (30%)

Multivariate Option 1
1. Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries.

```{r}
subset2 <- gapminder %>%
  filter(country %in% c("Chile", "Argentina", "Peru")) %>%
  select(country, year, lifeExp, gdpPercap)

subset2_wide <- subset2  %>%
   pivot_wider(id_cols     = year, 
              names_from  = country, 
              names_sep   = "_", 
              values_from = c(lifeExp, gdpPercap)) 

subset2_wide %>%
knitr::kable(align = "c", caption = "Life Expectancy & GDP percapita by Year Selected Countries")

```


2. Re-lengthen the data.

```{r}
(subset2_wide %>% 
  pivot_longer(cols      = c(-year), 
               names_to  = c(".value", "country"), 
               names_sep = "_")) %>%
knitr::kable(align = "c", caption = "Life Expectancy & GDP percapita by Year Selected Countries")
```




## Exercise 3: Table Joins (30%)

```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")

guest %>%
knitr::kable(align = "c", caption = "Guest List")
email %>%
knitr::kable(align = "c", caption = "Email List")
```


## 3.1 (10%)
 1. For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.

 * First I have to fix the guess names in the email file and rename the variable guest to name:
```{r}
email_list <- email %>% 
  separate_rows(guest, sep = ", ") %>% 
   rename(name = guest)
```

 * Then I can merge the two files: guest and emails, with guest names as identifier
```{r}
guest %>% 
  left_join(email_list, by = "name")
```
 


## 3.2 (10%)
Who do we have emails for, yet are not on the guestlist?
```{r}
email_list %>%
  anti_join(guest, by = "name")
```


## 3.3 (10%)
Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).


















