---
title: "hw02"
author: "JMF"
date: "27/09/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(gapminder)
library(tidyverse)
library(dplyr)
library(magrittr)
library(qwraps2)
args(summary_table)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r gapminder, echo = FALSE}
#summary table for Life Exp.


summary_lifeexp <-
  list("Life Expentancy" =
         list("min" = ~ min(gapminder$lifeExp),
              "max" = ~ max(gapminder$lifeExp),
              "mean" = ~ mean(gapminder$lifeExp),
              "sd" = ~ sd(gapminder$lifeExp)),
       "Years" = 
         list("min" = ~ min(gapminder$year),
              "max" = ~ max(gapminder$year)
              ))

table <- summary_table(gapminder, summary_lifeexp)
#Overall summary
table
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

