---
title: "hw02"
author: "JMF"
date: "27/09/2019"
output:
  #pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(gapminder)
library(tidyverse)
library(dplyr)
library(magrittr)
library(qwraps2)
library(ggplot2)
options(qwraps2_markup = 'markdown') # default is late
args(summary_table)
```

## 1.1 and 1.2
Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.
Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.
```{r , echo = TRUE}
subset <- filter(gapminder, country %in% 
                   c("Canada", "Mexico", "United States"), 
                 year > 1969,  year < 1981)
subset %>% select(country,gdpPercap)
```
## 1.3 
Filter gapminder to all entries that have experienced a drop in life expectancy. Be sure to include a new variable that’s the increase in life expectancy in your tibble. Hint: you might find the lag() or diff() functions useful.

```{r , echo = TRUE}
le_drop <- select(gapminder, country, year, continent, lifeExp) %>%
  group_by(country) %>%
  mutate(le_delta = lifeExp - lag(lifeExp))
  
  filter(le_drop, le_delta<0)
```

## 1.4a 
Max GDP per capita experienced by each country.
Filter gapminder so that it shows the max GDP per capita experienced by each country.
```{r , echo = TRUE}
  gapminder %>%
  select(country, year, continent, gdpPercap) %>%
  arrange(country) %>%
  group_by(country) %>%
  top_n(1, wt = desc(gdpPercap)) ## gets the max

```

## 1.5 
Scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2

```{r , echo = TRUE}
filter(gapminder, country %in% c("Canada")) %>%
ggplot(aes(x=log(gdpPercap),y=lifeExp)) + geom_point() 
```

## 2 
Explore individual variables with dplyr

```{r gapminder, echo = TRUE}
#summary table for Life Exp.

summary_lifeexp <-
  list("Life Expentancy" =
         list("min" = ~ min(gapminder$lifeExp),
              "max" = ~ max(gapminder$lifeExp),
              "mean" = ~ mean(gapminder$lifeExp),
              "sd" = ~ sd(gapminder$lifeExp)),
       "Years" = 
         list("min" = ~ min(gapminder$year),
              "max" = ~ max(gapminder$year)
              ))


summary_table(gapminder, summary_lifeexp)

```

## 2a

Grouped by Continent
```{r , echo = TRUE}

summary_table(dplyr::group_by(gapminder, continent), summary_lifeexp)

```




## 3

Average Life expectancy in time by continent

```{r , echo = F}
gapminder %>%
  group_by(continent, year) %>%
  summarise(lifeExp=mean(lifeExp)) %>%
  ggplot(aes(x=year, y=lifeExp, color=continent)) +
  geom_line(size=1) + 
  geom_point(size=1.5) +
  geom_smooth(method="loess")
```



## 3
Life expectancy of each country in time by continent

```{r , echo = F}
gapminder %>%
ggplot(aes(year, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  theme(legend.position = 'none')
```


## 3
Frequency of contries by continent

```{r , echo = F}
gapminder %>%
ggplot(aes(x=continent, fill=continent)) + 
  geom_bar(aes(y=..count../12)) +
  labs(y="Number of countries") +
  guides(fill=FALSE)
```


## 4

This is the correct code. Here %in% instead of == makes the filter to take all elements in country that match the countries Rwanda and Afghanistan

```{r , echo = T}
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))

```

