---
title: 'STAT 545A Assignment 03: dplyr/ggplot2 Part II'
output:
  #pdf_document: default
  html_document: default
---
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(gapminder)
library(gridExtra)
library(grid)
library(spelling)
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Pick three of the six tasks below, and produce:

- a tibble, using dplyr as your data manipulation tool;
- an accompanying plot of data from the tibble, using ggplot2 as your visualization tool; and
- some dialogue about what your tables/figures show (doesn’t have to be much).


## The tasks I picked and preamble:

In this assignment I used the gapminder dataset which contains country data from 1952 to 2007 (12 years) for 142 countries in 5 continents. Specifically, I focused on the variables
 - GDP percapita (US$, inflation-adjusted) from the World Bank (WB) and
 - Life Expectancy (years), defined as as the average number of years a newborn child would live if
current mortality patterns were to stay the same.

## Task Option 2
Get the maximum and minimum of GDP per capita for all continents.

In the table below, I summarized the minimun and maximun GDP percapita observed in the database by continent. Africa shows the lowest GDP percapita and Asia the highest.

```{r}
rangeGdpP <- gapminder %>% 
  group_by(continent) %>% 
  summarize(min_gdpPercap = min(gdpPercap),
            max_gdpPercap = max(gdpPercap),
            mean_gdpPercap = mean(gdpPercap))
rangeGdpP
```


In the figure below, I create a graph that contains the range for each continent. I included also the average GDP percapita as well. Asia is the continent that shows the wider range of values. 

```{r}
ggplot(rangeGdpP) + 
  geom_errorbar(aes(x=continent, ymin=min_gdpPercap, ymax=max_gdpPercap, color=continent, width=.5))  + 
  geom_point(aes(x=continent, y=mean_gdpPercap)) +
  scale_y_log10() + 
  ylab("GDP percapita") + xlab("Continent")  
```




## Task Option 3
Look at the spread of GDP per capita within the continents.

The table below depicts the variation within each continent about the GDP per capita. As seen in the previous task, Asia presents the higher variation with a standar deviation considerably higher than the rest of the continents.

```{r}
gapminder %>% 
group_by(continent) %>% 
summarize(mean=mean(gdpPercap),
          min=min(gdpPercap), 
          max=max(gdpPercap), 
          std=sd(gdpPercap), 
          q25=quantile(gdpPercap,0.25), 
          q50=quantile(gdpPercap, 0.5), 
          q75=quantile(gdpPercap, 0.75))
```


The figure below, shows the de

```{r}
ggplot(gapminder, aes(x=continent, y=gdpPercap)) +
  geom_boxplot(alpha=0.5, size=1, shape=1,)  + geom_jitter(width = 0.2, aes(colour = year)) +
  scale_y_log10()
```


# Task Option 4
Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
```{r}

```

# Task Option 5
How is life expectancy changing over time on different continents?
```{r}
gapminder %>%
    group_by(continent, year) %>%
    summarise(lifeExp=median(lifeExp)) 
```




```{r}
ggplot(gapminder, 
       aes(year,lifeExp, colour = continent)) + 
  geom_smooth() + ggtitle("Life Expextancy over Time by Continent") +
  xlab("Year") + ylab("Avg. Life Expectancy") 
```

# Task Option 6
Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.
```{r}

```

# Ex

```{r}
df <- gapminder %>% 
  # for each continent
  group_by(continent) %>%  
  summarize(max = max(gdpPercap), min = min(gdpPercap))
```

```{r}
dfplt <- df %>% 
  ggplot(aes(continent)) + 
  # plot the points in the same graph and differentiate by color
  geom_point(aes(y = max, colour = "max")) + 
  geom_point(aes(y = min, colour = "min")) +  
  ggtitle("The max and min of GDP per capita for all continents")
# display the table and graph side by side
grid.arrange(tableGrob(df), dfplt, nrow = 1)
```




```{r, echo = TRUE}
ex <- gapminder %>% 
group_by(continent) %>% 
summarize(mean=mean(gdpPercap),
          std=sd(gdpPercap))

exg <- ex %>%
  ggplot(aes(continent)) +
  geom_point(aes(y = mean, colour = "mean")) + 
  geom_point(aes(y = std, colour = "std")) +  
  ggtitle("continents")

  grid.arrange(tableGrob(ex), exg, nrow = 1)

```



