---
title: "Assignment 1"
author: "JMF"
date: "16/09/2019"
bibliography: biblio.bib
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A look into greenhouse gas emissions concentrations

As Solomon2018 

In this document, I explore the concentrations of greenhouse gas emissions using data from the Data Science Lab (dslab). The data measures the concentrations of the three main greenhouse gases carbon dioxide, methane and nitrous oxide. The data was collected from the Law Dome Ice Core in Antarctica. Selected measurements are provided every 20 years from 1 to 2000 CE [@MacFarling2006]

```{r cars}
summary(cars)
```

```{r, echo=FALSE}
library("dslabs")
data("greenhouse_gases")
co2=greenhouse_gases[,1:3]
summary(greenhouse_gases$concentration)
```


## Concentration of CO2 emissions in time

In figure 1 (refered this) we can observe the increase of the concentrations of CO2 emissions in the last 2000 years. 

```{r, echo=FALSE}
library("dslabs")
library(dplyr)
library(ggplot2)
data("greenhouse_gases")
greenhouse_gases_co2 <- greenhouse_gases %>%
  filter(gas == "CO2")
ggplot(greenhouse_gases_co2, aes(x = year, y = concentration)) +
  geom_point() + ggtitle("Concentration of CO2") +
  xlab("Year") + ylab("ppm") + theme(
    plot.title = element_text(hjust = 0.5))
```



```{r, echo=FALSE}
library("dslabs")
library(dplyr)
library(ggplot2)
data("greenhouse_gases")
greenhouse_gases_co2 <- greenhouse_gases %>%
  filter(gas == "CO2", year>1800)
ggplot(greenhouse_gases_co2, aes(x = year, y = concentration)) +
  geom_point() + ggtitle("Concentration of CO2") +
  xlab("Year") + ylab("ppm") + theme(
    plot.title = element_text(hjust = 0.5))
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
