---
title: "A look into greenhouse gas emissions"
author: "Juan Fercovic"
date: "16/09/2019"
bibliography: biblio.bib
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview of the presentation

- Introduction
- Summary of the data
- CO2 emissions in time
- Adjustment of emissions in the last 200 years

## Introduction

- Concentrations of CO2 collected from the Law Dome Ice Core in Antarctica [@MacFarling2006].
- CO2 concentrations measured in parts per million (ppm).
- Years 0 to 2000.



## Summary of the data

```{r echo=FALSE, message=FALSE, warning=FALSE}
library("dslabs")
data("greenhouse_gases")
library(dplyr)
greenhouse_gases_co2 <- greenhouse_gases %>%
  filter(gas == "CO2")
summary(greenhouse_gases_co2)
```

## CO2 emissions in time

```{r, echo=FALSE}
library("dslabs")
library(ggplot2)
data("greenhouse_gases")
greenhouse_gases_co2 <- greenhouse_gases %>%
  filter(gas == "CO2")
ggplot(greenhouse_gases_co2, aes(x = year, y = concentration)) +
  geom_point() + ggtitle("Concentration of CO2") +
  xlab("Year") + ylab("ppm") + theme(
    plot.title = element_text(hjust = 0.5))
```


## Adjustment of emissions in the last 200 years

```{r, echo=FALSE}
greenhouse_gases_co2 <- greenhouse_gases %>%
  filter(gas == "CO2", year>1800)
lnconc <-  log(greenhouse_gases_co2$concentration)
lnyr <-  log(greenhouse_gases_co2$year)
scatter.smooth(x=greenhouse_gases_co2$year, y=lnconc , ylab="log of ppm of CO2",
               xlab="Year", main="CO2 emissions since 1800") 
  

```

## References:
